\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-07-03 paper/subdocuments/methods.Rnw
minor changes methods.Rnw

\subsection{Outline of applied methods}

@@

\subsubsection{Relative distribution (RD) methods }

@@

% Probability Density function as a base for the RD
The goal of RD is to study the differences between two distributions. A common example could be the income distribution for man and women. Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point. To describe what we are going to compare, we start with defining the two distributions. One represents the reference population $Y_{0}$ and the other the comparison population $Y$. $Y$ represents our measure of interest (income). A first visual approach is to compare the two probability density functions (PDF). The PDF is a function $f(y)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $y$. Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%. Out of the comparison of the PDF, it is possible to see, which values of $y$ are more and less probable. This allows already to spot distributional differences over the whole scale of $Y$ visually. \\

The PDF can be characterized by its cumulative distribution function (CDF). The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $y$, which is the probability that a randomly chosen value is less than or equal to $y$. The relative distribution of $Y$ to $Y_{0}$ is then defined as

>>>>

\subsection{Outline of research strategy and applied methods}

@@

In the last section we described the advantages and drawback of tax-data discussing five aspects, which we regard as crucial concerning the assessment of inequality. To get a feeling of the importance of these aspects, we exploit the FTA-tax data as far as possible and perform several insightful calculations addressing four of the five mentioned aspects. No further investigation is possible regarding aspect (1) concepts on measuring economic resource. But for the other four thematic areas, we can provide deeper insight. In general our main strategy is to apply different possible concepts within one conceptional area (income, units, measurement, population coverage) while holding other conceptual differences constant. With this strategy we want to show, the sensitiveness the assessment of inequality is especially if one looks at time trends.\\

To fulfill the above described task, we use two techniques. To assess the development over time, we calculate in general gini-coefficients for all possible time points, allowing us the make time trends visible and to compare the result out of the FTA-tax data to other publications. For selected periods we expand the analysis with relative distribution methods, which allows an in-depth analysis of distributional differences and therefore catches the shortcomings of gini-coefficients. \\

To calculate Gini-coefficients we follow common definitions, like it is described for example in \citep{jann_einfuhrung_2005}. Because the relative distribution methods are not that widely know, we provide a short introduction of the important concepts in this section. We additional expand the method section with an extended analysis about the importance of non-taxed, because the issue of non-taxed is a especially thorny one.

\subsubsection{Relative distribution (RD) framework}

@@

% Probability Density function as a base for the RD
The goal of RD is to study the differences between two distributions. A common example could be the income distribution for man and women. Subject of the comparison can also be two distributions describing the same population, but stemming from two different data sources like survey or tax data or even comparison of the same source/population but for different time points, like the income distribution out of tax data for Switzerland today compared to an earlier time point. To describe how the two distributions are gone be transformed into a relative distribution, we start with defining the two distributions. One represents the reference population $Y_{0}$ and the other the comparison population $Y$. $x$ represents our measure of interest (income). A first visual approach is to compare the two probability density functions (PDF). The PDF is a function $f(x)$ which describes the distribution of probability over the outcome set and is defined for all possible values of $x$. Also this function integrates to 1, which means that the sum of all probabilities over all possible values is 100\%. Out of the comparison of the PDF, it is possible to see, which values of $x$ are more and less probable. This allows already to spot distributional differences over the whole scale of $Y$ visually. \\

The PDF can be characterized by its cumulative distribution function (CDF). The CDF can be formulated as $F(y)$, which represents the cumulative probability at point $x$, which is the probability that a randomly chosen value is less than or equal to $x$. The relative distribution of $Y$ to $Y_{0}$ is then defined as
